Watch Out for the Bully! Job Interference Study on Dragonfly Network

### 2016-03-30 draft.tex
made changes for section 1, 2, 3

\subsection{Dragonfly Network}
\label{sec:network}
The dragonfly is a two-level hierarchical topology. It consists of several groups connected by all-to-all links. In each group, there could be ``a" number of routers connected via all-to-all local channels. For each router, there could be ``p" number of compute nodes attached to it via terminal links. The router has ``h" number of global channels that used as intergroup connections through which the router connected to other routers in different groups. Thus, the radix of each router is $k = a+h+p-1$. Different computing centers could choose different values for ``a", ``p" and ``h" when deploying their dragonfly network. The adoption of proper $a$, $h$, $p$ involves many factors, such as system scale, building cost and workload.

It is recommended that for the load balance purpose, a proper dragonfly configuration should follow $a=2p=2h$ \cite{kim-micro}. Under this configuration, the total number of groups ``g" in the dragonfly network would be $g = a*h+1 $, the total number of compute nodes ``N" in the network would be $N = p*a*g $. An implementation of dragonfly network follows such configuration is illustrated in Figure \ref{fig:dragonfly-overview}. There are 6 routers in each group, each router has 3 compute nodes attached to it. This dragonfly implementation consists of 19 groups and 342 nodes in total.

\begin{figure}[h!] 
  \centering
  \includegraphics[width=0.48\textwidth]{dragonfly-overview}
  \caption{Dragonfly Network Overview. There are 19 groups, 6 routers in each group, 3 compute nodes connected to each router, 342 nodes in total. Job $J1$ requires for 7 nodes, gets its allocation by random placement. Job $J2$ requires for 8 nodes, gets its allocation by contiguous placement. }
  \label{fig:dragonfly-overview}
\end{figure}

@@

For a parallel application requiring multiple compute nodes, job placement policy refers to the way of assigning the tasks of the application onto the system by a system software such as batch scheduler \cite{xu-cluster14}. In this work, we study two alternative placement policies considered by the community for dragonfly systems:

\textbf{Contiguous Placement:} In this policy, the compute nodes are assigned to job in a consecutive manner. The assignment fills up a group first, only crosses group boundaries and start to fill the next group as necessary. Shown in Figure \ref{fig:dragonfly-overview}, $J2$ gets the required amount of nodes by contiguous placement. Contiguous placement confines the tasks of a job into the same group, which may result in local network congestion and increase the possibility of hot-spots.

\textbf{Random Placement:} In this policy, job will get a set of nodes that randomly selected from all the available nodes in the system. The nodes assigned to $J1$ in Figure \ref{fig:dragonfly-overview} would be attached to different routers in different groups. Routers may be shared by different jobs when random placement is in use. Random placement can distribute the tasks of a job uniformly across the whole system, thus avoid the possible local congestion and reach load balance. However, random placement may cause possible congestion on global links.

\subsection{Routing on Dragonfly}
\label{sec:routing-schemes}
The routing policy refers to the strategy adopted to route a message (a stream of packets) from the source router to the destination router. Previously studied routing policies for dragonfly network include minimal routing, adaptive routing \cite{dally-dragonfly}, progressive adaptive routing\cite{jiang} and some of their variations \cite{won-prog-adaptive}. In this work we study three widely adopted routing policies on dragonfly networks.

\textbf{Minimal:} In this policy, a message from source router takes the the shortest path to the destination router. When there are multiple shortest paths between the source and destination router, the message will be evenly divided among the paths. Minimal routing can guarantee the minimum hops the message takes from source to destination. However, it may result in congestion along the shortest path.

\textbf{Adaptive:} In this policy, the path a message takes will be adaptively chosen between shortest and non-shortest path, depending on the congestion situation along that path. For the non-shortest path, an intermediate router will be randomly chosen. The message first takes the shortest path from source to that intermediate router, then it takes the shortest path from that intermediate router to the destination. Since routing path is chosen adaptively based on the congestion situation, adaptive routing can avoid local congestion and possible hot-spots. Adaptive routing collapses to minimal routing in the absence of traffic congestion.

\textbf{Progressive Adaptive:} In this policy, the decision to route minimally at each hop in the source group will be re-evaluated. Any decision to route non-minimally at source router or at a subsequent hop is permanent and won't be re-evaluated \cite{jiang}. The progressive adaptive is capable of handling the case where a global link is congested but this information has not yet propagated back to the source router \cite{jiang}. The packet will take the non-minimal route only when the congestion is encountered.

@@

It is difficult to experiment with concurrently running jobs on HPC system \cite{zhou-ipdps-2015}\cite{jain-sc14}\cite{bhatele-sc11}\cite{jokanovic-ipdps-2015}. One reason is that job placement and routing policies are part of system configuration, it is impossible for user to make changes at will. Another reason is that it is unrealistic to reserve the system exclusively to run the same batch of jobs with desired placement and routing configurations and compare the results. Last but not least, configurable dragonfly networks that allow us to perform the exploration presented in this paper are hard to come by for the time being. Therefore, we resort to simulation in our work.

\subsection{Simulation Tool}
\label{sec:simulation-tool}
A simulation toolkit named CODES, Co-Design of Multilayer Exascale Storage Architectures, enables the exploration of simulating different HPC networks with high fidelity and great scalability \cite{codes}. CODES supports dragonfly network\cite{codes-dragonfly} \cite{misbah-tpds}, as well as other topologies such as torus\cite{misbah-pads-2014}\cite{ning-pads-2011} and fat-tree\cite{ning-pads-2015}, with packet-level high-fidelity simulation. It can drive these models through an MPI simulation later that utilizing DUMPI traces\cite{sst}. In this work, we perform an in-depth analysis of network performance as well as job interference when different job placement and routing policies are in use on dragonfly network.

@@

We choose three representative parallel applications from the DOE Design Forward Project\cite{designforwardwebpage}. We believe these applications can be representative of a wide array of applications running on leadership-class supercomputers. Specifically, we study the \emph{Algebraic MultiGrid Solver} (AMG), \emph{Geometric Multigrid V-Cycle from production elliptic solver} (MultiGrid) and \emph{Crystal Router MiniApp} (CrystalRouter).

\textbf{AMG:} The Algebraic MultiGrid Solver is a parallel algebraic multi-grid solver for linear systems arising from problems on unstructured mesh physics packages. It has been derived directly from the BoomerAMG solver that is being developed in the Center for Applied Scientific Computing (CASC) at LLNL \cite{amg}.

\textbf{MultiGrid:} The geometric multi-grid v-cycle from the production elliptic solver BoxLib is a software framework for massively parallel block-structured adaptive mesh refinement (AMR) codes \cite{boxlib}. It is widely used for structured grid physics packages.

\textbf{CrystalRouter:} The CrystalRouter MiniApp is the communication kernel of the full application Nek5000 \cite{nek5000}, which is a spectral element CFD application developed at Argonne National Laboratory. It features spectral element multi-grid solvers coupled with a highly scalable, parallel coarse-grid solver that is widely used for projects including ocean current modeling, thermal hydraulics of reactor cores, and spatiotemporal chaos.

@@

The parameters for building the dragonfly network been studied in our work are inspired the model proposed by Kim et. al \cite{kim-micro}. Our dragonfly network consists of 33 groups, each group contains 8 routers, and each router has 4 compute nodes attached to it. Overall, there are 264 routers and 1056 compute nodes in the network. The aggregated bandwidth of terminal link,  local and global channels are proportional to those of Cray Cascade system\cite{faanes}. In this work, we study network performance as well as job interference when 6 different job placement and routing policy combinations are in use on the dragonfly network, which are summarized in Table \ref{tab: placement routing configs}.\footnote{With respect to random placement, we experiment with 50 sets of distinctive allocation generated by random placement. The corresponding experimental results are median chosen, which intended to eliminate the possibility of variation.}

@@

Our analysis focus on the following metrics:
\begin{itemize}

    \item \textbf{Network Traffic:} The traffic refers to the amount of data go through each port of the router. For each router, the traffic go through its terminal link, local channel and global channel are analyzed respectively.

    \item \textbf{Network Saturated Time:} The saturated time refers to the time period when the buffer of certain port is full. Saturated time can denote the congestion on certain port.

    \item \textbf{Communication Time:} The communication time of each MPI rank refers to the time spent by the rank to complete its message exchange with all other ranks. For each job, this metric refers to the time spent to complete the message exchange by all its ranks.

@@

Since our study focus on network performance and job interference during communication, the time spent in computation by each MPI rank will not be counted in our experiments.

@@

Two sets of parallel workloads are used in this work. Workload \Rmnum{1} consists of AMG, MultiGrid and CrystalRouter. As we see in Table \ref{tab:apps-detail}, AMG has the least amount of data transfer, making it the least communication-intensive job in Workload \Rmnum{1}. Workload \Rmnum{2} consists of sAMG, MultiGrid and CrystalRouter. sAMG, a synthetic version of AMG, is generated by increasing AMG data transfer amount by 100x. sAMG is the most communication-intensive job in Workload \Rmnum{2}.

Workload \Rmnum{1} is used to identify the ``bully" behavior between concurrently running jobs on dragonfly network. Workload \Rmnum{2} is used to verify the results and conclusion obtained from the study about Workload \Rmnum{1}.

@@

\section{Study of Parallel Workload \Rmnum{1}}
%\section{Identify the ``Bully"}
\label{sec:workload-1}

We study the network performance as well as job interference from two perspectives. First, we analyze the network performance when Workload \Rmnum{1} is running under different placement and routing configurations. Then, we isolate each job from the workload, and analyze the interference by scrutinizing the traffic go through the routers that serving each job. These analysis allow us to identify the ``bully" in the workload.

@@

\begin{table}[ht]
\begin{center}
\caption{Average time spent on communication by all MPI ranks when Workload I is running on dragonfly network under 6 different placement and routing configurations.} 
\label{tab:wkld-commtime}
\begin{tabular}{l c c c c c c }
\toprule % Top horizontal line
\toprule
&\multicolumn{6}{c}{Placement and Routing Configurations} \\ 
\cmidrule(l){2-7}
	      & CM & CA & CPA & RM & RA & RPA \\ % Column names row
\midrule % In-table horizontal line
Time(ms)  & 476  & 300  & 301  & 255  & 265  & 264  \\ % Content row 1

>>>>

\subsection{Dragonfly Network}
\label{sec:network}
The dragonfly is a two-level hierarchical topology, consisting of several groups connected by all-to-all links. In each group, there could be $a$ routers connected via all-to-all local channels. For each router, there could be $p$ compute nodes attached to it via terminal links. The router has $h$ global channels used as intergroup connections through which the router connected to other routers in different groups. Thus, the radix of each router is $k = a+h+p-1$. Different computing centers could choose different values for  $a$, $h$, $p$ when deploying their dragonfly network. The adoption of proper $a$, $h$, $p$ involves many factors, such as system scale, building cost and workload.

It is recommended that for the load balance purpose, a proper dragonfly configuration should follow $a=2p=2h$~\cite{kim-micro}. Under this configuration, the total number of groups, denoted as $g$ in the dragonfly network would be $g = a*h+1 $, the total number of compute nodes denoted as $N$ in the network would be $N = p*a*g $. An implementation of dragonfly network follows such configuration is illustrated in Figure~\ref{fig:dragonfly-overview}. There are six routers in each group, each router has three compute nodes attached to it. This dragonfly network consists of 19 groups and 342 nodes in total.

\begin{figure}[h!] 
  \centering
  \includegraphics[width=0.48\textwidth]{dragonfly-overview}
  \caption{Dragonfly Network Overview. There are 19 groups, six routers in each group, three compute nodes connected to each router, 342 nodes in total. Job $J1$ requires seven nodes, getting its allocation by random placement. Job $J2$ requires eight nodes, getting its allocation by contiguous placement.}
  \label{fig:dragonfly-overview}
\end{figure}

@@

For a parallel application requiring multiple compute nodes, job placement policy refers to the way of assigning tasks of the application onto the system by a system software such as batch scheduler~\cite{xu-cluster14}. In this work, we study two alternative placement policies considered by the community for dragonfly systems:

\textbf{Contiguous Placement:} In this policy, the compute nodes are assigned to a job in a consecutive manner. The assignment first fills up a group, then crosses group boundaries and starts to fill the next group if necessary. As shown in Figure~\ref{fig:dragonfly-overview}, $J2$ gets eight nodes by contiguous placement. Contiguous placement confines the tasks of a job into the same group, which may result in local network congestion and increase the possibility of hot-spots.

\textbf{Random Placement:} In this policy, a job gets a set of nodes randomly selected from all the available nodes in the system. The nodes assigned to $J1$ in Figure~\ref{fig:dragonfly-overview} are attached to different routers in different groups. Routers may be shared by different jobs when random placement is in use. Random placement can distribute the tasks of a job uniformly across the whole system to avoid the possible local congestion. However, random placement may cause possible congestion on global links.

\subsection{Routing on Dragonfly}
\label{sec:routing-schemes}

The routing policy refers to the strategy adopted to route a message (a stream of packets) from the source router to the destination router. Previously studied routing policies for dragonfly network include minimal routing, adaptive routing~\cite{dally-dragonfly}, progressive adaptive routing~\cite{jiang} and some of their variations~\cite{won-prog-adaptive}. In this work we study three widely adopted routing policies on dragonfly networks.

\textbf{Minimal:} In this policy, a message takes the shortest path from the source router to the destination router. When there are multiple shortest paths between the source and destination router, the message will be evenly divided among the paths. Minimal routing can guarantee the minimum hops the message takes from source to destination. However, it may result in congestion along the shortest path.

\textbf{Adaptive:} In this policy, the path a message takes will be adaptively chosen between shortest and non-shortest paths, depending on the congestion situation along those paths. For the non-shortest path, an intermediate router will be randomly chosen. The message takes the intermediate router as a via point, connecting the source and destination router through two separate shortest paths. Adaptive routing can avoid hot-spots in the presence of congestion and collapse to minimal routing otherwise.

\textbf{Progressive Adaptive:} In this policy, the decision to route minimally at each hop in the source group will be re-evaluated. Any decision to route non-minimally at source router or at a subsequent hop is permanent and will not be re-evaluated~\cite{jiang}. The progressive adaptive is capable of handling the case where a global channel is congested but the source router has not been informed yet~\cite{jiang}. The packet will take the non-minimal route only when the congestion is encountered.

@@

It is difficult to experiment with concurrently running jobs on HPC system~\cite{zhou-ipdps-2015}\cite{jain-sc14}\cite{bhatele-sc11}\cite{jokanovic-ipdps-2015}. One reason is that job placement and routing policies are part of system configuration, which is impossible for user to make changes at will. Another reason is that it is unrealistic to reserve the system exclusively to run the same batch of jobs with desired placement and routing configurations and compare the results. The third reason is that configurable dragonfly networks that allow us to perform the exploration presented in this paper are hard to come by for the time being. Therefore, we resort to simulation in our work.

\subsection{Simulation Tool}
\label{sec:simulation-tool}

A simulation toolkit named CODES, Co-Design of Multilayer Exascale Storage Architectures, enables the exploration of simulating different HPC networks with high fidelity and great scalability~\cite{codes}. CODES supports dragonfly~\cite{codes-dragonfly}~\cite{misbah-tpds}, torus~\cite{misbah-pads-2014}~\cite{ning-pads-2011}, and fat-tree~\cite{ning-pads-2015} networks with packet-level high-fidelity simulation. It can drive these models through an MPI simulation layer utilizing DUMPI traces~\cite{sst}. In this work, we perform an in-depth analysis of network performance as well as job interference when different job placement and routing policies are in use on the dragonfly network.

@@

We choose three parallel applications from the DOE Design Forward Project~\cite{designforwardwebpage}, which can represent of a wide array of applications running on leadership-class supercomputers. Specifically, we study the \emph{Algebraic MultiGrid Solver} (AMG), \emph{Geometric Multigrid V-Cycle from Production Elliptic Solver} (MultiGrid) and \emph{Crystal Router MiniApp} (CrystalRouter).

\textbf{AMG:} The Algebraic MultiGrid Solver is a parallel algebraic multi-grid solver for linear systems arising from problems on unstructured mesh physics packages. It has been derived directly from the BoomerAMG solver developed in the Center for Applied Scientific Computing (CASC) at LLNL~\cite{amg}.

\textbf{MultiGrid:} The geometric multi-grid v-cycle from the production elliptic solver BoxLib is a software framework for massively parallel block-structured adaptive mesh refinement (AMR) codes~\cite{boxlib}, which are used for structured grid physics packages.

\textbf{CrystalRouter:} The Crystal Router MiniApp is the communication kernel of the full application Nek5000 \cite{nek5000}, a spectral element CFD application developed at Argonne National Laboratory. It features spectral element multi-grid solvers coupled with a highly scalable, parallel coarse-grid solver that is widely used for projects including ocean current modeling, thermal hydraulics of reactor cores, and spatiotemporal chaos.

@@

The parameters for building the dragonfly network studied in our work are inspired by the model proposed by Kim et. al~\cite{kim-micro}. Our dragonfly network consists of 33 groups, each of which contains 8 routers. Each router has four compute nodes attached to it. Overall, there are 264 routers and 1056 compute nodes in the network. The aggregated bandwidth of terminal link,  local and global channels are proportional to those of the Cray Cascade system~\cite{faanes}. In this work, we study both the network performance and job interference when six different job placement and routing policy combinations are in use on the dragonfly network, which are summarized in Table~\ref{tab: placement routing configs}.

%\footnote{With respect to random placement, we experiment with 50 sets of distinctive allocation generated by random placement. The corresponding experimental results are median chosen, which intended to eliminate the possibility of variation.}

@@

Our analysis focuses on the following metrics:
\begin{itemize}

    \item \textbf{Network Traffic:} The traffic refers to the amount of data going through each router. We analyze the traffic on the terminal link, local and global channel of each router.

    \item \textbf{Network Saturated Time:} The saturated time refers to the time period when the buffer of a certain port in the router is full. We analyze the saturated time of ports corresponding to terminal link, local and global channel.

    \item \textbf{Communication Time:} The communication time of each MPI rank refers to the time it spends in completing all its message exchanges with other ranks. \footnote{Since our study focuses on network performance and job interference during communication, the computation time of each MPI rank will not be counted in our experiments.
}

@@

@@

Two sets of parallel workloads are used in this work. Workload~\Rmnum{1} consists of AMG, MultiGrid and CrystalRouter. In Table~\ref{tab:apps-detail}, AMG has the least amount of data transfer, making it the least communication-intensive job in Workload~\Rmnum{1}. Workload~\Rmnum{2} consists of sAMG, MultiGrid and CrystalRouter. sAMG, a synthetic version of AMG, is generated by increasing AMG data transfer amount by 100x. sAMG is the most communication-intensive job in Workload~\Rmnum{2}.

Workload~\Rmnum{1} is used to identify the ``bully" behavior among concurrently running jobs on the dragonfly network. Workload~\Rmnum{2} is used to verify the results and conclusion obtained from the study of Workload~\Rmnum{1}.

@@

\section{Study of Parallel Workload~\Rmnum{1}}
%\section{Identify the ``Bully"}
\label{sec:workload-1}

We study the network performance and job interference from two perspectives. Firstly, we analyze the network performance when Workload \Rmnum{1} is running under different placement and routing configurations. Secondly, we isolate each job from the workload, and analyze the interference by scrutinizing the traffic going through the routers that serve each job. The analysis allows us to identify the ``bully" in the workload.

@@

\begin{table}[ht]
\begin{center}
\caption{Average communication time by all MPI ranks when Workload I is running on the dragonfly network under six different placement and routing configurations.} 
\label{tab:wkld-commtime}
\begin{tabular}{l c c c c c c }
\toprule % Top horizontal line
\toprule
&\multicolumn{6}{c}{Placement and Routing Configurations} \\ 
\cmidrule(l){2-7}
	      & CM & CA & CPA & RM & RA & RPA \\ % Column names row
\midrule % In-table horizontal line
Time(ms)  & 476  & 300  & 301  & 255  & 265  & 264  \\ % Content row 1
